### 2025-04-26 thesis/thesis.tex
Query synthesis

>>>>

\subsubsection{Query Synthesis}

\cite{SQLizer} introduced SQLizer, which synthesizes SQL queries from natural language \citep{SQLizer}. This paper
presents a novel approach when it comes to NL2SQL as it is merging prevalent semantic parsing techniques (outlined above)
with an program synthesis (or query synthesis) approach. SQLizer makes use of a three stage processing model for 
natural language models: first generating a sketch of the query using semantic parsing, then using type-directed
synthesis to complete the sketch and finally using automated repair, if required.

\citeauthor*{SQLizer} show that alternating between quantitative repair and synthesis yields results that beat
state-of-the-art approaches like NaLIR. SQLizer is fully automated and database-agnostic, requiring no knowledge
of the underlying schema. The authors evaluated SQLizer on 455 queries across three databases, where it ranked the
correct query in the top 5 results for about 90\% of cases, significantly outperforming NaLIR \citep{NALIR},
the previous state-of-the-art system \citep{SQLizer}.

Potential short commings of this approch include queries which yield empty results, dealing with language variations
as SQLizer is still using semantic parsing, and domain-specific terminology, all while still requiring users
to select from multiple query options which reduces the overall usability of the system \citep[p.22-23]{SQLizer}.
